The effect of starvation on the dynamics of consumer populations

### 2016-07-13 Manuscript/Yeakel_Kempes_Redner-7-10-2016.tex
new stuff

The behavioral ecology of most, if not all, organisms is influenced by the energetic state of individuals.
An individual's energetic state directly influences how it invests its stores in an uncertain environment.
Such behaviors are generally manifested as trade-offs, which often concern investing in individual maintenance and growth (somatic effort) or allocating energy towards reproduction (reproductive effort) \citep{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}. %among a host of other behavioral duties [REFS].
The timing of these behaviors is often important and is under strong selective pressure, as they tend to have large effects on the future fitness of the organism \citep{Mangel:1988uaa}.
%For example, rotifers that reproduce have lower survival rates than those that don't when they are under nutritional stress.
To what extent, and when, organisms invest in somatic vs. reproductive expenditures may be driven by habitat, seasonality, evolutionary history, inter- or intra-specific interactions, as well as resource limitation.
Importantly, the influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to repeated delays or shifts in reproduction over the course of an organism's life. %and the subsequent loss of nutritional reserves
%Such tradeoffs are also bound to have population-level consequences, although this has received little theoretical or empirical attention.

@@

\noindent {\bf Allometric constraints} \\ \nonumber
%[Link Allometry stuff to our model - what does it provide to the story]
%[Introduce biological and linked constraints]
%[Allometries capture vast amounts of diversity via a single parameter --- body size]
%[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
Chris - Linking paragraph here -- what does allometry provide to the story? Biological, linked constraints... vast amounts of diversity in a small number of parameters. Etc.

@@

The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues tissue from the starving state to the full state.
We again use the balance given in Equation \ref{balance} to find the timescale to return to the mature mass from a given reduced starvation mass. The general solution to Equation \ref{balance} is given by
\begin{equation}
m\left(t\right)=\left[1-\left(1-\frac{b}{a}m_{0}^{1-\alpha}\right)e^{-b\left(1-\alpha\right)t}\right]^{1/\left(1-\alpha\right)}\left(\frac{a}{b}\right)^{1/\left(1-\alpha\right)}
\end{equation}
with $a=B_{0}/E_{m}$ and $b=B_{m}/E_{m}$. We are then interested in the timescale, $t_{\rho}=t_{2}-t_{1}$, which is the time it takes to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, which has the final form of 
\begin{equation}
t_{\rho}=\frac{\log \left(1-\left(M \left(\frac{a}{b}\right)^{\frac{1}{\alpha -1}}\right)^{1-\alpha }\right)-\log \left(1-\left(M \epsilon  \left(\frac{a}{b}\right)^{\frac{1}{\alpha
   -1}}\right)^{1-\alpha }\right)}{(\alpha -1) b}.
\end{equation}

@@

>>>>

The behavioral ecology of most, if not all, organisms is influenced by the energetic state of individuals.
An individual's energetic state directly influences how it invests its stores in an uncertain environment.
Such behaviors are generally manifested as trade-offs, which often concern investing in individual maintenance and growth (somatic effort) or allocating energy towards reproduction (reproductive effort) \citep{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}. %among a host of other behavioral duties [REFS].
The timing of these behaviors is often important and is under strong selective pressure, as they tend to have large effects on the future fitness of the organism \citep{Mangel:1988uaa}.
%For example, rotifers that reproduce have lower survival rates than those that don't when they are under nutritional stress.
To what extent, and when, organisms invest in somatic vs. reproductive expenditures may be driven by habitat, seasonality, evolutionary history, inter- or intra-specific interactions, the distribution of resources.
Importantly, the influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to repeated delays or shifts in reproduction over the course of an organism's life. %and the subsequent loss of nutritional reserves
%Such tradeoffs are also bound to have population-level consequences, although this has received little theoretical or empirical attention.

@@

\noindent {\bf Allometric constraints} \\ \nonumber
%[Link Allometry stuff to our model - what does it provide to the story]
%[Introduce biological and linked constraints]
%[Allometries capture vast amounts of diversity via a single parameter --- body size]
%[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
Our model explores a wide variety of possible dynamics yet we have not described the realistic regimes occupied by organisms where the challenge is to constrain the covariation of rates. Allometric scaling relationships highlight common constraints and average trends across large ranges in body size and species diversity. Many of these relationships can be derived from a small set of assumptions and below we describe a framework for the covariation of timescales and rates across the range of mammals for each of the key parameters of our model. We are able to define the regime of dynamics occupied by the entire class of mammals along with the key differences between the largest and smallest mammals.

@@

The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues tissue from the starving state to the full state.
We again use the balance given in Equation \ref{balance} to find the timescale to return to the mature mass from a given reduced starvation mass. The general solution to Equation \ref{balance} is given by
\begin{equation}
m\left(t\right)=\left[1-\left(1-\frac{b}{a}m_{0}^{1-\eta}\right)e^{-b\left(1-\eta\right)t}\right]^{1/\left(1-\eta\right)}\left(\frac{a}{b}\right)^{1/\left(1-\eta\right)}
\end{equation}
with $a=B_{0}/E_{m}$ and $b=B_{m}/E_{m}$. We are then interested in the timescale, $t_{\rho}=t_{2}-t_{1}$, which is the time it takes to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, which has the final form of 
\begin{equation}
t_{\rho}=\frac{\log \left(1-\left(M \left(\frac{a}{b}\right)^{\frac{1}{\eta -1}}\right)^{1-\eta }\right)-\log \left(1-\left(M \epsilon  \left(\frac{a}{b}\right)^{\frac{1}{\eta
   -1}}\right)^{1-\alpha }\right)}{(\eta -1) b}.
\end{equation}

@@

 \begin{figure}[h]
 	\centering
 	\includegraphics[width=0.75\textwidth]{Growth-trajectory-diagram.pdf}
 	\caption{
 	}
 	\label{growth-diagram}
 \end{figure}
